---
title: "p8105_hw6_yx2954"
author: "Yiran Xu"
date: "2024-11-28"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(p8105.datasets)
```

# Problem 1 

Load data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Create and apply bootstrap generating function

```{r}
boot_sample <- function(df) {
  sample_frac(df, replace = TRUE)
}
```

We should also do a quick check to see if this is working.
```{r bootstrap_sample_check, fig.cap="bootstrap_sample_check"}
boot_sample(weather_df) |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(alpha = .5) +
  stat_smooth(method = "lm")
```

That looks about right. Then we generate 5,000 bootstrap samples
```{r}
boot_straps = tibble(strap_number = 1:5000) |>
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = weather_df))
  )
```

```{r}
bootstrap_results <- boot_straps |>
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin, data = df)),
    r2 = map_dbl(models, \(model) broom::glance(model)$r.squared),
    log_beta0_beta1 = map_dbl(models, \(model) {
      coefs <- broom::tidy(model)$estimate
      log(coefs[1] * coefs[2])
    })
  ) |>
  select(-strap_sample, -models) |>
  unnest(log_beta0_beta1) 
```

Plot the distribution of two sample estimates
```{r estimate_plot, fig.cap="estimate_plot"}
bootstrap_longer = bootstrap_results |>
  pivot_longer(
    cols = c(r2, log_beta0_beta1),
    names_to = "statistic",
    values_to = "value"
  )

estimate_plot = bootstrap_longer |>
  ggplot(aes(x = value)) +
  geom_density(alpha = 0.5, fill = "blue") +
  facet_wrap(~statistic, scales = "free", labeller = labeller(
    statistic = c(
      r2 = "R^2",
      log_beta0_beta1 = "log(β0 * β1)"
    )
  )) +
  labs(
    title = "Bootstrap Distributions",
    x = "Value",
    y = "Density"
  )

print(estimate_plot)
```

Get numeric values of sample estimates
```{r}
bootstrap_summary <- bootstrap_results %>%
  summarize(
    r2_mean = mean(r2),
    r2_sd = sd(r2),
    log_beta0_beta1_mean = mean(log_beta0_beta1),
    log_beta0_beta1_sd = sd(log_beta0_beta1)
  )
```

The plot shows that both sample estimates are nearly normally distributed. The sample mean of log(beta0 * beta1) is 2.013649 and standard error is 0.02389236. The maen of r2 is 0.9114209 and the standard error is 0.00849257.

Get CI for both sample estimates
```{r}
ci_r2 = quantile(bootstrap_results$r2, c(0.025, 0.975))
ci_log_beta0_beta1 <- quantile(bootstrap_results$log_beta0_beta1, c(0.025, 0.975))
```

The 95% CI of r2 is (0.8940791, 0.9271204), the 95% CI of log(beta0 * beta1) is (1.964630, 2.058959)

# Problem 2

Load data and data cleaning
```{r}
homicide_df <- read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") 

homicide_df = homicide_df |>
  mutate(city_state = paste(city, state, sep = ", "),
         status = if_else(disposition == "Closed by arrest", 1, 0),
         victim_age = as.numeric(victim_age)
           ) |>  
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")) |>
  select(-city, - state, -disposition)
```

Logistic regression in Baltimore, MD

```{r}
baltimore_df = homicide_df |>
  filter(city_state == "Baltimore, MD")

baltimore_glm = glm(status ~ victim_age + victim_sex + victim_race,
                     data = baltimore_df,
                     family = binomial()) |>
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)

baltimore_or = baltimore_glm |>
  filter(term == "victim_sexMale") |>
  select(term, estimate, conf.low, conf.high)

print(baltimore_or)
```
The Odd ratio for solving homicides comparing male victims to female victims is 0.4255117, with a 95% CI of (0.3241908, 0.5575508). It suggests that the result is significant as the higher boundary is smaller than 1, and the probability of solving homicides of male victims is 42.6% as likely as that of female victims.


Run glm for each of the cities

```{r}
all_city_results = homicide_df |>
  group_by(city_state) |>
  nest() |>
  mutate(
    glm_model = map(data, ~ glm(status ~ victim_age + victim_sex + victim_race, data = ., family = binomial())),
    tidy_model = map(glm_model, ~ broom::tidy(.x, exponentiate = TRUE, conf.int = TRUE))
  ) |>
  unnest(tidy_model) |>
  filter(term == "victim_sexMale") %>%
  select(city_state, estimate, conf.low, conf.high) |>
  arrange(estimate)
```

Create a plot that shows the estimated ORs and CIs for each city.
```{r all_city_plot, fig.cap="all_city_plot", fig.height=7, fig.width=10}
all_city_plot = all_city_results |>
  ggplot(aes(x = reorder(city_state, estimate), y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip() +
  labs(
    title = "Adjusted Odds Ratios for Solving Homicides (Male vs Female)",
    x = "City",
    y = "Odds Ratio",
  ) +
  theme_minimal()

print(all_city_plot)
```
